Stability and Chaos in Multi Agent Reinforcement Learning

### 2020-07-24 Documents/ChaoticDynamics/AAAI Submission/main.tex
minor mods

% The file aaai20.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\setlength\titlebox{2.5in} % If your paper contains an overfull \vbox too high warning at the beginning of the document, use this
% command to correct it. You may not alter the value below 2.5 in
\title{Stability and Chaos in Multi Agent Reinforcement Learning}
%Your title must be in mixed case, not sentence case. 
% That means all verbs (including short verbs like be, is, using,and go), 
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\author{Aamal Hussain, Francesco Belardinelli\textsuperscript{\rm 1}\thanks{Something to do with the
CDT}\\ % All authors must be in the same font size and format. Use \Large and \textbf to achieve this result when breaking a line
\textsuperscript{\rm 1}Imperial College London\\ %If you have multiple authors and multiple
% affiliations
% use superscripts in text and roman font to identify them. For example, Sunil Issar,\textsuperscript{\rm 2} J. Scott Penberthy\textsuperscript{\rm 3} George Ferguson,\textsuperscript{\rm 4} Hans Guesgen\textsuperscript{\rm 5}. Note that the comma should be placed BEFORE the superscript for optimum readability
Do we give the address?\\
publications20@aaai.org % email address must be in roman text type, not monospace or sans serif
}
 \begin{document}

@@

@@

In this section we will establish some of the preliminaries which are required to follow the subsequent sections. The first is 'Dynamical Behaviours' which considers the evolution of agent strategies as they learn on an iterated game. This study aims to classify, based on the agent parameters and the payoff matrices, which of these behaviours will be seen. The second is 'Dynamics of Q-Learning'. These are a set of equations which model the aforementioned strategy evolution. It is on these dynamics that we will perform our analysis.

@@

    It has long been established that, upon lifting the strong
    assumptions made by traditional game theory (such as the
    rationality of agents and complete information), player
    strategies can result in much more complex behaviour than
    convergence to a Nash Equilibrium (NE). In fact, this is even true
    on what would commonly be regarded as 'simple' games such as
    tic-tac-toe and prisoner's dilemma \cite{Galla2011,
      Sato2002}. These behaviours include: convergence to a unique
    equilibrium (though not always to an NE), convergence to one of
    multiple equilibria, limit cycles and chaos. These behaviours are
    shown in Figure \ref{fig::DynamicalBehaviours}. Of these, the most
    preferable is, of course, convergence to a unique equilibrium,
    although it is still possible to study systems with multiple
    equilibria or limit cycles \cite{Strogatz2000}. However, it would
    be difficult to control systems whose dynamics are governed by
    chaos (though research into controlling chaos is ongoing and rife
    with opportunity \cite{Fradkov2009}) and so MARL techniques should
    avoid this. It would, therefore, be a useful endeavour to
    determine the conditions under which these sorts of behaviours
    arise.

    \begin{figure}[h]
        \centering
        \includegraphics[width=1.1\textwidth]{Figures/DynamicalBehaviours}
        \caption{ \label{fig::DynamicalBehaviours} Different types of dynamical behaviour
       displayed
        by learning agents. Here the x-axis is the probability with which agent 1 chooses a given action, whilst the y-axis denotes the same probability for agent 2. a) Figure drawn from \cite{Tuyls2006AnGames}.
        Convergence
        to a unique fixed point in the upper right corner (1, 1). This fixed point is unique, in
        that all trajectories, regardless of initialisation, will converge to this point. b) Limit Cycle, the trajectories converge to cyclic behaviour c,
        d) Chaotic behaviour, here small deviatons in the initial conditions can grow
        exponentially. b-d drawn from \cite{Sanders2018}}
    \end{figure}

@@

    Here, $\alpha$ and $\tau$ are the parameters of the agent; Sanders et al. refer to these as the
    memory and intensity of choice parameters respectively. Agent 1 takes action $i$ with probability
    $x_i$ while Agent 2 takes action $j$ with probability $y_j$. If these actions are taken, the agents
    receive payoff $a_{ij}$ and $b_{ji}$ respectively. With these equations, it is possible to
    predict the expected behaviour of Q-Learning agents, as shown in Figure 
    \ref{fig::TuylsExperiments}.

@@

    It is clear, both from (\ref{eqn::EOM}) and Figure \ref{fig::TuylsExperiments}, that the long-term strategy selection of
    these agents is determined by the parameters $\alpha, \tau$ and the payoffs $a_{ij}, b_{ij}$. We
    then pose the question: how do these elements influence the types of behaviours seen during
    learning on an iterated game? In other words, under what parameter selections are we likely to
    see convergence to unique equilibria, multiple equilibria, limit cycles or chaos?

@@

@@

To verify our theoretical results, and to examine the underlying structure of stability and chaos in
Multi-Agent Q Learning, we perform a series of numerical experiments by varying the parameters $\Gamma$ and
$\alpha$ whilst keeping $\tau$ fixed. The purpose of this experiment was to determine a heuristic estimate for the Liapunov exponent, which determines whether initial conditions which start close to one another will converge or diverge from each other. The former indicates a stable system (in the sense of Liapunov) whilst the latter signals instability.

@@

    \caption{$\tau = 0.05$, $\gamma = 0.75$, $\alpha \in [0, 1]$, $\Gamma \in [-1, 1]$. Each
    simulation is run for 1 $\times 10^4$ iterations and the Liapunov exponent is determined. For each
    combination of $\alpha$, $\Gamma$ the game is played 100 times, each with random payoff matrices
    and initial conditions. The average liapunov exponent is determined and shown in the heatmap.
    \label{fig::NumericalExperiments}}
\end{figure}

@@

>>>>

% The file aaai20.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\setlength\titlebox{2.5in} % If your paper contains an overfull \vbox too high warning at the beginning of the document, use this
% command to correct it. You may not alter the value below 2.5 in
\title{Stability and Chaos in Multi Agent Reinforcement Learning}
%Your title must be in mixed case, not sentence case. 
% That means all verbs (including short verbs like be, is, using,and go), 
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\author{Aamal Hussain, Francesco Belardinelli\textsuperscript{\rm 1}\thanks{Something to do with the
CDT}\\ % All authors must be in the same font size and format. Use \Large and \textbf to achieve this result when breaking a line
\textsuperscript{\rm 1}Imperial College London\\ %If you have multiple authors and multiple
% affiliations
% use superscripts in text and roman font to identify them. For example, Sunil Issar,\textsuperscript{\rm 2} J. Scott Penberthy\textsuperscript{\rm 3} George Ferguson,\textsuperscript{\rm 4} Hans Guesgen\textsuperscript{\rm 5}. Note that the comma should be placed BEFORE the superscript for optimum readability
Do we give the address? \fb{normally the submission is blind.}\\
publications20@aaai.org % email address must be in roman text type, not monospace or sans serif
}

\input{macros}

 \begin{document}

@@

\fb{possibly add research question below.}

\subsection{Related Work}

\fb{feel free to add the Related Work}

@@

In this section we will establish some of the preliminaries which are
required to follow the subsequent sections. The first is 'Dynamical
Behaviours' \fb{maybe italics?} which considers the evolution of agent strategies as they
learn on an iterated game. This study aims to classify, based on the
agent parameters and the payoff matrices, which of these behaviours
will be seen. The second is 'Dynamics of Q-Learning'. These are a set
of equations which model the aforementioned strategy evolution. It is
on these dynamics that we will perform our analysis.

@@

    It has long been established \fb{add refs?} that, upon lifting the strong
    assumptions made by traditional game theory (such as the
    rationality of agents and complete information), player
    strategies can result in much more complex behaviour than
    convergence to a Nash Equilibrium (NE). In fact, this is even true
    on what would commonly be regarded as 'simple' games such as
    tic-tac-toe and prisoner's dilemma \cite{Galla2011,
      Sato2002}. These behaviours include: convergence to a unique
    equilibrium (though not always to an NE), convergence to one of
    multiple equilibria, limit cycles and chaos. These behaviours are
    shown in Figure \ref{fig::DynamicalBehaviours}. Of these, the most
    preferable is, of course, convergence to a unique equilibrium,
    although it is still possible to study systems with multiple
    equilibria or limit cycles \cite{Strogatz2000}. However, it would
    be difficult to control systems whose dynamics are governed by
    chaos (though research into controlling chaos is ongoing and rife
    with opportunity \cite{Fradkov2009}) and so MARL techniques should
    avoid this. It would, therefore, be a useful endeavour to
    determine the conditions under which these sorts of behaviours
    arise.

    \begin{figure}[h]
        \centering
        \includegraphics[width=1.1\textwidth]{Figures/DynamicalBehaviours}
        \caption{ \label{fig::DynamicalBehaviours} Different types of dynamical behaviour
       displayed
        by learning agents. Here the x-axis is the probability with which agent 1 chooses a given action, whilst the y-axis denotes the same probability for agent 2. a) Figure drawn from \cite{Tuyls2006AnGames}.
        Convergence
        to a unique fixed point in the upper right corner (1, 1). This fixed point is unique, in
        that all trajectories, regardless of initialisation, will converge to this point. b) Limit Cycle, the trajectories converge to cyclic behaviour c,
        d) Chaotic behaviour, here small deviatons in the initial conditions can grow
        exponentially. b-d drawn from \cite{Sanders2018}}
\fb{one figure from a reference is fine, but two maybe are not necessary.}
    \end{figure}

@@

    Here, $\alpha$ and $\tau$ are the parameters of the agent; Sanders et al. refer to these as the
    memory and intensity of choice parameters respectively \fb{what are $\alpha$ and $\tau$ exactly?}. Agent 1 takes action $i$ with probability
    $x_i$ while Agent 2 takes action $j$ with probability $y_j$. If these actions are taken, the agents
    receive payoff $a_{ij}$ and $b_{ji}$ respectively. With these equations, it is possible to
    predict the expected behaviour of Q-Learning agents, as shown in Figure 
    \ref{fig::TuylsExperiments} \fb{discuss limitations as well}.

@@

    It is clear, both from (\ref{eqn::EOM}) and Figure \ref{fig::TuylsExperiments}, that the long-term strategy selection of
    these agents is determined by the parameters $\alpha, \tau$ and the payoffs $a_{ij}, b_{ij}$. We
    then pose the question: how do these elements influence the types of behaviours seen during
    learning on an iterated game? In other words, under what parameter selections are we likely to
    see convergence to unique equilibria, multiple equilibria, limit cycles or chaos? \fb{state this explicitly as a research question, possibly in the contribution.}

@@

\fb{what will be the content of this section? is it the derivations?}

@@

To verify our theoretical results, and to examine the underlying structure of stability and chaos in
Multi-Agent Q Learning, we perform a series of numerical experiments by varying the parameters $\Gamma$ and
$\alpha$ whilst keeping $\tau$ fixed. The purpose of this experiment was to determine a heuristic estimate for the Lyapunov exponent, which determines whether initial conditions which start close to one another will converge or diverge from each other. The former indicates a stable system (in the sense of Lyapunov) whilst the latter signals instability \fb{add refs to all of this?}.

@@

    \caption{$\tau = 0.05$, $\gamma = 0.75$, $\alpha \in [0, 1]$, $\Gamma \in [-1, 1]$. Each
    simulation is run for 1 $\times 10^4$ iterations and the Lyapunov exponent is determined. For each
    combination of $\alpha$, $\Gamma$ the game is played 100 times, each with random payoff matrices
    and initial conditions. The average Lyapunov exponent is determined and shown in the heatmap.
    \label{fig::NumericalExperiments}}
\end{figure}

@@

\fb{Liapunov $\to$ Lyapunov}
